Four decades of water quality change in the upper San Francisco Estuary

### 2017-01-13 sftrends_manu.Rnw
working on flo obs text comparison

A comparison of trends with flow-normalized components relative to observed data is justified because flow and nutrient concentrations were linked at many of the stations in the study area, similar to \cref{fig:dinc10}.  These comparisons are made relative to changes in the magnitude, significance, and direction of the trend, all of which have important implications for decision-making. Overall, the differences in apparent trends underscore the critical importance of considering flow effects in decision-making, where flow is affected by complex relationships between inflow, water withdrawals, and exchange with the larger Bay.

>>>>

<<echo = F, cache = F>>=
data(trnds_seasyr)

cmp <- data.frame(trnds_obs, dat = 'obs') %>% 
  rbind(., data.frame(trnds_nrm, dat = 'nrm')) %>% 
  filter(cat %in% c('1976-1995', '1996-2013')) %>% 
  select(dat, Site_Code, resvar, cat, ztest, perchg)

zcmp <- select(cmp, -perchg) %>% 
  mutate(ztest = ifelse(ztest < .05, 'sig', 'ns')) %>% 
  spread(dat, ztest) %>% 
  mutate(
    sigtons = ifelse(obs == 'sig' & nrm == 'ns', 1, 0), 
    nstosig = ifelse(obs == 'ns' & nrm == 'sig', 1, 0)
  )

chgcmp <- select(cmp, -ztest)# %>% 
  # mutate(ztest = ifelse(ztest < .05, 'sig', 'ns')) %>% 
  # spread(dat, ztest) %>% 
  # mutate(
  #   sigtons = ifelse(obs == 'sig' & nrm == 'ns', 1, 0), 
  #   nstosig = ifelse(obs == 'ns' & nrm == 'sig', 1, 0)
  # )

A comparison of trends with flow-normalized results from \ac{wrtds} relative to observed data is justified because flow and nutrient concentrations were linked at many of the stations in the study area, similar to \cref{fig:dinc10}.  These comparisons are made relative to changes in the magnitude, significance, and direction of trends, all of which have important implications for decision-making. For all \Sexpr{eng(nrow(zcmp))} trend comparisons in \cref{fig:trndcomp1} regardless of site, nitrogen analyte, and time period, \Sexpr{eng(sum(zcmp$nstosig))} comparisons had trends that were insignificant with the observed data but significant with flow-normalized results. The magnitude of all comparisons changed

Overall, the differences in apparent trends underscore the critical importance of considering flow effects in the interpretation of environmental changes, where flow in the Delta is affected by complex relationships between river inputs, water withdrawals, and tidal exchange with the larger Bay.

### 2017-01-13 sftrends_manu.tex
working on flo obs text comparison

A comparison of trends with flow-normalized components relative to observed data is justified because flow and nutrient concentrations were linked at many of the stations in the study area, similar to \cref{fig:dinc10}.  These comparisons are made relative to changes in the magnitude, significance, and direction of the trend, all of which have important implications for decision-making. Overall, the differences in apparent trends underscore the critical importance of considering flow effects in decision-making, where flow is affected by complex relationships between inflow, water withdrawals, and exchange with the larger Bay.

>>>>

A comparison of trends with flow-normalized results from \ac{wrtds} relative to observed data is justified because flow and nutrient concentrations were linked at many of the stations in the study area, similar to \cref{fig:dinc10}.  These comparisons are made relative to changes in the magnitude, significance, and direction of trends, all of which have important implications for decision-making. For all sixty trend comparisons in \cref{fig:trndcomp1} regardless of site, nitrogen analyte, and time period, thirteen comparisons had trends that were insignificant with the observed data but significant with flow-normalized results. The magnitude of all comparisons changed

Overall, the differences in apparent trends underscore the critical importance of considering flow effects in the interpretation of environmental changes, where flow in the Delta is affected by complex relationships between river inputs, water withdrawals, and tidal exchange with the larger Bay.
